Hysteresis in human computation:\\ how one task affects another

### 2014-12-17 docs/publications/2014.09_SCIENCE/man.tex
major revisions based on Dereks v1 critiques

\author
{Edward Newell, Derek Ruths,\\
\\
\normalsize{\texttt{edward.newell@mail.mcgill.ca}}\\
\normalsize{\texttt{druths@networkdynamics.org}}\\
\normalsize{School of computer science, McGill University,}\\
\normalsize{3630 rue University, Montreal, Quebec, H3A 0C6, Canada}\\
\\
}

@@

Faced with the scale and complexity of the data revolution, researchers and
industry are increasingly relying on \textit{human computation}.  Microtask
platforms, like Amazon Mechanical Turk, act like a human compute server,
making it possible to access human cognition quickly and inexpensively. 
This has lead to broad adoption across a spectrum of disciplines
in the humanities and computer science, as well as in research into medical 
imaging technologies and other applications where expert opinion would 
normally be sought.  Here we show that an inherent
feature of microtask platforms, the repetition of similar tasks,
acts as a strong, but as yet unexplored source of bias.  
These \textit{intertask effects} rival the
severity of overtly framing a task's purpose.  
Using image-labeling as a canonical microtask, we found that prior 
tasks influence the vocabulary that a worker uses when labeling 
subsequent images in both profound an nuanced ways.  Prior tasks not only 
alter what a worker focuses on, 
but also modulate the richness and specialization of her vocabulary.  
While intertask effects can be a source of systematic bias, our
results suggest that, when properly controlled, they can be leveraged
to hone worker focus and specificity.  Though we focused on microtask 
platforms due to their emerging prominence, our results apply to  
citizen science, crowdsourcing, and any application involving repetitive 
tasks relying on human judgment.
\end{sciabstract}

\section*{Introduction}
Microtask platforms are online marketplaces on which \textit{requesters} 
post batches of tasks, and \textit{workers} complete them
for remuneration, a sense of purpose, and for fun
\cite{kazai2013analysis,Antin20122925}.  
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results 
\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.

@@

Unlike CPUs and GPUs, the experiments we present here suggest HPUs
are subject to \textit{hysteresis}, meaning that their outputs depend
on the history of their inputs.  Given the
increasing prevalence of microtask-based research methods, other factors 
affecting HPU outputs have received considerable attention.
This includes various task-design parameters such as the level of 
pay \cite{kazai2013analysis}, the training \cite{le2010ensuring} and 
pre-screening \cite{paolacci2010running} of workers, and user-interface 
design factors \cite{Finnerty2013}.  
Researchers have also investigated \textit{framing} effects, 
such as the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or altering the problem description
\cite{thibodeau2013natural}.  Results in psychology show that people are 
susceptible to various kinds of priming 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, 
task-repetition effects \cite{Gass1999549,sohn2001task}.  But to our 
knowledge, no study has investigated the practical ramifications 
of intertask effects on human computation.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  Image-labeling tasks are among
the most common MTurk tasks \cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, a trend which will probably continue 
in the long term because of their role in computer vision research 
\cite{5543192}.
In our experiments, workers performed a set of \textit{initial tasks}, 
followed by a set of \textit{test tasks}. Crucially, no distinction was
made between these sets from the perspective of the worker. We varied the 
initial tasks (for example, having workers label images of food or objects), 
and observed the ensuing effects on responses in test tasks.  
As a point of 
comparison, we also included treatments in which we framed the tasks in terms 
of different purposes.

The results showed that intertask effects are stronger than framing effects.  
Framing effects were comparable to intertask effects only when the workers 
were asked to actively reiterate the frame by means of a combo-box input.
Using the wordnet corpus (\#), we found that intertask 
effects can hone a worker's focus, and boost the specialization
and lexical richness of her labels.  Thus, while intertask effects may be a 
serious source of bias, they can also be leveraged to help achieve
expert-level performance in crowdsourcing, when properly controlled.

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.6]{figs/images.pdf}
	\caption{
		Examples of images used for various experiments and treatments:
		A) initial task for \textit{task1:food}, B) initial task for 
		\textit{task1:obj}, C) test task for treatments in \textit{exp1}, 
		D) initial task for \textit{task2:food}, 
		E) initial task for \textit{task2:cult},  
		F) test task for treatments in \textit{exp2}.  
		The full set of experimental materials is shown in the 
		supplementary text.
	}

@@

We performed two experiments, \textit{exp1} and \textit{exp2}, which consisted
of various pairs of treatments.  The treatments in each pair were subjected to
the same exposure modality, either initial tasks or framing, but differed in
the content of the exposure.  Table \ref{table:treatments}  
summarizes the treatments and introduces a naming convention that we use 
in what follows.
In \textit{exp1}, workers in the treatment pair \textit{task1} initially
labeled 5 images containing either food (treatment \textit{task1:food}), or 
objects (\textit{task1:obj}), and then performed test tasks, in 
which they 
labeled images containing both food and objects (see Fig.~\ref{fig:task} for 
examples of the images used). In \textit{exp1}, we included two other 
treatment pairs, \textit{frame1} and \textit{echo1}, in which workers did not 
perform initial tasks, but instead were given information which framed the 
purpose of the tasks.  Workers in \textit{frame1} were told that the tasks 
were ``Funded by the laboratory for the visual perception of 
\{ Food and Ingredients $\vert$ Objects and Tools~\}''.  In \textit{echo1},
we sharpened the frame, by indicating that ``The purpose of this study is to 
understand the visual perception of 
\{ Food and Ingredients $\vert$ Objects and Tools~\}'' and then asking the
worker to reiterate the purpose of the task using a combo-box input.

In contrast to \textit{exp1}, which was based on concepts of food and
objects, \textit{exp2} was based food and \textit{culture}.
Since culture is more abstract, one might expect different effects 
in \textit{exp2}.  In \textit{exp2}, we again investigated both
intertask effects (in treatment pair \textit{task2}) and framing effects 
(\textit{frame2}), but with a slight variation in the framing treatments.
In \textit{frame2}, workers were not only exposed to a frame, but also 
performed a set of initial tasks before the test tasks.  Since its purpose 
was still
to investigate framing effects, the initial tasks in \textit{frame2}
were not varied. Thus, \textit{frame2} provides insight into the moderating 
effect that intervening tasks might have on framing effects.

In all treatments, workers were shown images one at a time, and provided five 
labels for each.  From the perspective of the worker, there was no
distinction or interruption between the initial and test tasks.  We used
workers' unique identifiers to ensure that they only participated in
one treatment of a single experiment.

\setlength{\tabcolsep}{2pt}
\begin{table}[t]
\centering
\begin{tabular}{ c c c c c }
		\hline \noalign{\smallskip}
		\multicolumn{3}{c}{\textbf{Treatment name}} & \parbox[c]{1.4cm}{\centering \textbf{Frame}} & \parbox[c]{1.3cm}{\centering \textbf{Initial\\ tasks}}	\\

		\noalign{\smallskip} \hline \noalign{\smallskip}

		\multirow{6}{*}{ \parbox[c]{0.8cm}{ \phantom{XX} exp1}} 
			& \multirow{2}{*}{task1} & food & none & food\\
			& & obj & none & objects\\

			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{frame1} & food & food & none\\
			& & obj & objects & none\\

			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{echo1} & food & food & none\\
			& & obj & objects & none\\

		\noalign{\smallskip} \hline \noalign{\smallskip}

		\multirow{4}{*}{\parbox[c]{0.8cm}{exp2}} 
			& \multirow{2}{*}{task2}  &  food & none & food\\
			& 	&  cult & none & culture\\
			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
			& \multirow{2}{*}{frame2} & food & food & food\\
			& 	& cult & culture & food\\

		\noalign{\smallskip} \hline  
	\end{tabular}

	\caption{ 
		We use the treatment names shown to refer to specific sections of 
		the experimental data.
	}
	\label{table:treatments}
\end{table}

\paragraph*{Measuring computational hysteresis.}
While our study builds on research into the psychological mechanisms
of task performance, there is a crucial difference in our present goal. 
Our goal is to measure the practical ramifications of intertask effects on 
the microtask methodology. Whereas a statistically significant effect, even 
if small, might be sufficient to distinguish one psychological mechanism 
from another (\#), in our study, 
\textit{effect size} is central.  Thus our main goal is not merely to reject 
the null hypothesis that initial tasks have no effect
(we do this in the supplementary text).

To couch intertask effects in a practical setting, we imagine
that, after a worker performs a task, her response is used to make some 
kind of yes-or-no decision, $\mathcal{D}$.  For example, this decision 
could be whether or not to flag a certain comment
below a news article as inappropriate; but for now, we leave 
$\mathcal{D}$ unspecified.  Since workers' responses may
vary, repeating the same task might lead to different outcomes for 
$\mathcal{D}$.  This decision not only models the reality that responses are
actually used---introducing $\mathcal{D}$ allows us to define \textit{bias}
in a practical yet general way. We define bias in terms of the
change in the probability of a positive outcome from $\mathcal{D}$, induced 
by prior exposures.

The kind of decision that $\mathcal{D}$ represents will affect our 
measurement of bias.  For instance, if $\mathcal{D}$ is random, or if 
it always generates ``yes'' 
irrespective of the worker's response, then of course we will 
find zero bias.  Likewise, there is an upper limit to the bias induced on $\mathcal{D}$. In a precise sense, the bias of $\mathcal{D}$ can only be as
great as the bias inherent in the worker's response 
(this is formalized in the supplementary text).  Therefore, to estimate the 
bias induced in workers' outputs, we adopt an 
adversarial approach: we seek to construct the decision that 
is the most susceptible to workers' biases. In the supplementary text, we 
show that this can be formulated as a  machine learning problem, wherein one
constructs a classifier that infers the prior exposure of a worker from her 
responses in test tasks.  The bias, $\theta$, induced by 
prior exposures, can then be bounded using the classifier's accuracy, 
$\eta$ (\#):
\begin{equation}
	\theta \geq 2\eta - 1
	\label{l1}
\end{equation}

\paragraph{The strength and dynamics of intertask effects.}
For each treatment pair, we trained a na\"ive Bayes classifier (\# in the 
supplementary material we discuss the choice of classifier, the
approach to measuring accuracy, and draw a comparison to the results obtained 
if a support vector machine is used.) to distinguish workers from one 
treatment (e.g \textit{task1:food}) from those of the conjugate treatment 
(e.g. \textit{task1:obj}).  In \textit{task1} and \textit{task2}, intertask 
effects profoundly biased worker outputs, by approximately 30 and 50\% 
respectively, whereas we could not detect 
a statistically significant bias induced by framing in \textit{frame1} and 
\textit{frame2} (see Fig.~\ref{fig:theta}A).  The bias induced
by framing was only appreciable in \textit{echo1}, reaching approximately 
35\%.  
We suspect that asking workers to reiterate the
frame, as was done in \textit{echo1}, might be interpreted by workers as
a direct signal that the frame should be acted upon.  In that sense, 
\textit{echo1} is perhaps more indicative of the effects of an 
\textit{instruction}.  These observations show that
prior tasks profoundly influence the responses of workers, producing an
effect that is stronger than framing, and on par with the combination of 
framing and reiteration.

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Bias as measured by a na\"ive Bayes classifier, $\theta_\mathrm{NB}$,
		induced in image-labeling tasks, due to framing and 
		intertask effects. 
		A) Bias in the responses to five test tasks,
		measured between the pairs of treatments indicated on the abscissa.  
		B) Bias between the treatments of \textit{task1}, 
		measured for individual test tasks, and plotted as a function of 
		task position.  Error bars show the 95\% confidence intervals.
	}
	\label{fig:theta}
\end{figure}

To observe the dynamics of intertask effects, we trained a na\"ive Bayes
classifier to distinguish workers in \textit{task1} based on responses to
individual test tasks.  We permuted 
the test tasks to eliminate effects arising from differences in the images 
therein (more details are provided in the supplementary text). The first test 
task shows the greatest bias, after which the effects drop somewhat, but 
substantial bias persists through the fifth test task
(see Fig. \ref{fig:theta}B).  Presumably, this decay profile reflects the
fact that workers are also influenced by their exposure to the test tasks, 
which begins to wash out the effects of initial tasks.

\paragraph{Detailed influences on worker outputs.} 
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/vocab_specificity.pdf}
	\caption{
		Intertask and framing effects on the vocabulary workers use to refer 
		to food(\#), when labeling images, for various treatment pairs.
		A) Relative change in the number references to food. 
		B) Relative change in the number of \textit{unique} 
		references to food (lexical richness). 
		C) Relative specialization of food references, based on the 
		hypernym-hyponym relations (this calculation is shown in the 
		supplementary text).  For all
		plots, a positive bar height indicates that the \textit{*:food}
		treatment has more of the quantity than its conjugate treatment.
	}
	\label{fig:specificity}
\end{figure}

Using the wordnet corpus, we analyzed exposure effects in finer detail.
Wordnet encodes semantic information in various ways, including in the form 
of synsets (groups of synonymous words), and hypernym-hyponym relations.
Hypernyms are generalizations (``bread'' is a hypernym of ``pumpernickel''),
while hyponyms are specialisations.

First, we used wordnet to identify nouns depicting foods (\# We 
operationalized food-references by taking all hyponyms of the synsets 
\texttt{food.n.01} and \texttt{food.n.02}, with the addition of ethnic foods 
learned by crawling a recipe website, as described in the supplementary text),
to see whether 
workers from the various food-exposed treatments (\textit{*:food}), 
provided more references to food than their conjugate treatments, overall.
The \textit{exp1:*:food} treatments
did produce more food references, but, unexpectedly, 
the number of food references in \textit{task2:food} was
less than that in \textit{task2:cult} (see Fig.~\ref{fig:specificity}A). Also, the effects on the
number of food references was, overall, somewhat smaller than one might 
expect given the extent of bias observed in Fig.~\ref{fig:theta}A.  This means that task and frame exposures
induce other effects which are more subtle than a bulk shift in the number 
of references to the primed concept (food).

Next, we looked at the change in the frequency of occurrence of particular 
words.  Surprisingly, the word ``food'' was the \textit{most suppressed} word
for workers in \textit{*:food} across all treatment pairs (Table S1 shows the
top five most promoted and suppressed words for each treatment pair).
Initially, this seems at odds with the fact that (at least in \textit{exp1})
the \textit{*:food} treatments had more food references than their conjugate
treatments.  But taken together, these facts indicate that \textit{*:food}
workers are opting 
to use more specialized food-references in place of ``food''.  
Indeed, when we look at the total number of unique
words (lexical richness) used to refer to food, we found that 
a significantly richer lexicon was used by all \textit{*:food} treatments,
except in \textit{frame2} (see Fig.~\ref{fig:specificity}B).  This effect 
was especially pronounced in \textit{task1} and 
\textit{task2}.

Finally, we directly calculated the degree of specialization of food 
references using wordnet's hypernym-hyponym relations.  To determine the 
relative specialization between the words of two treatments, we computed 
the proportion of words in the first treatment that were hyponyms of the 
those in the 
second, less those in the first that were hypernyms of those in the second 
(this calculation is formalized in the supplementary text).  Except in the
case of \textit{frame2}, workers in \textit{*:food} treatments used 
substantially more specialized terms in reference to food than workers 
from the conjugate treatments.

\paragraph{Positive and negative priming.}
Taken together, the results of the experiments presented here can be 
explained through a combination of positive and negative priming.
Positive priming (usually simply ``priming'') occurs when a prior stimulus 
predisposes a person to give certain responses in an ensuing task, and
is often observed as an increase in the speed or accuracy of a response, or
the ability to recognize objects in a briefer, quieter, or noisier signal (\#).
Negative priming occurs when, after being repeatedly exposed to a stimulus 
that is perceived to be non-salient, the person begins to ignore that 
stimulus (\#).

Workers exposed to images containing food will be (positively) primed, 
as memories, concepts, and vocabulary related to food are activated.  
However, we argue that as the worker labels successive images containing food,
the basic 
fact that an image contains food will not appear to be salient, since it 
does not serve to distinguish one image from another.  Thus, the most 
generic references to that fact, such as the label ``food'', 
are suppressed, as by negative priming, while more specialized
references are activated.  Whether workers in \textit{*:food} treatments
produce an excess of food references overall depends in part on the balance 
of these factors.  The reduction in food references from workers in 
\textit{task2:food} relative to \textit{task2:cult}, seen in in 
Fig.~\ref{fig:specificity}A, may also be
related to the novelty that workers in 
\textit{task2:cult} experience when they first encounter food in the test 
tasks.

>>>>

\author
{Edward Newell \\ Derek Ruths\\
\\
\normalsize{\texttt{edward.newell@mail.mcgill.ca}}\\
\normalsize{\texttt{druths@networkdynamics.org}}\\
\normalsize{School of computer science, McGill University,}\\
\normalsize{3630 rue University, Montreal, Quebec, H3A 0C6, Canada}\\
\\
}

@@

Faced with the ever-growing scale and complexity of data, researchers and
industry are increasingly relying on \textit{human computation} as a means
to extract useful knowledge.  Microtask
platforms, like Amazon Mechanical Turk, act like a human compute server,
making it possible to access human cognition quickly and inexpensively. 
This has lead to broad adoption across a spectrum of disciplines
in the social sciences and computer science, as well as in research into 
medical imaging technologies and other applications where expert opinion 
would normally be sought.  Here we show that an inherent
feature of microtask platforms and most crowdsourcing settings---the 
repetition of similar tasks---acts as a strong, but as yet unexplored source 
of bias.  These \textit{intertask effects} rival the
severity of overtly framing a task's purpose.  
Using image-labeling as a canonical microtask, we show that prior 
tasks influence the content that workers focus on, and the richness and
specialization of their vocabulary.    
While intertask effects can be a source of systematic bias, our
results suggest that, when properly controlled, they might be leveraged
to hone worker focus and specificity, and could play a key role in designing 
tasks that elicit expert-level judgments.
\end{sciabstract}

\section*{Main text}
Microtask platforms are online marketplaces in which \textit{requesters} 
post batches of tasks, and \textit{workers} complete them
for remuneration, a sense of purpose, and for fun
\cite{kazai2013analysis,Antin20122925}.  
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results 
\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.

@@

Given the increasing prevalence of microtask-based research methods, there
is a desire to understand the factors affecting the reliability of HPU
outputs.  Researchers have investigated task-design parameters such as the 
level of pay \cite{kazai2013analysis}, the training \cite{le2010ensuring} and 
pre-screening \cite{paolacci2010running} of workers, and user-interface 
design factors \cite{Finnerty2013}.  
Researchers have also investigated \textit{framing} effects, 
such as the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or altering the problem description
\cite{thibodeau2013natural}.  Results in psychology show that people are 
susceptible to various kinds of priming 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, 
task-repetition effects \cite{Gass1999549,sohn2001task}.  
Yet to our knowledge, no study has investigated the practical ramifications 
of intertask effects on human computation. The experiments we 
present here show that, unlike CPUs and GPUs, HPUs are subject to 
\textit{hysteresis}, meaning that their outputs depend on the 
\textit{history} of their inputs.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  Image-labeling tasks are among
the most common MTurk tasks \cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}.  This trend will probably continue 
in the long term because of their role in computer vision research 
\td{find another application} \cite{5543192}.  In microtask platforms, it is typical for work to be divided 
into bundles; on MTurk, these bundles are called Human Intelligence Tasks 
(HITs). Because of this bundling pattern, even if workers perform only one 
HIT, the opportunity for intertask effects arises.  For clarity, 
we will use ``task'' to mean the smallest repeatable unit of work, which, 
in our study, consists of labeling a single image.

In our first experiment, which we will refer to as 
\textit{intertask-food-objects}, workers were either assigned to the 
\textit{food} or \textit{objects} treatment.  Workers in these treatments
performed a set of five \textit{initial tasks}, which involved labeling
images depicting either food or (non-food) objects, depending on the worker's 
treatment (see Fig.~\ref{fig:task}A and B).  Following the initial 
tasks, workers performed five \textit{test tasks}, which were identical for 
both treatments.  The test tasks contained images depicting a 
mixture of food and other objects (see Fig.~\ref{fig:task}C).  Five initial 
tasks, together with the test tasks, comprised one HIT 
(\# workers were only allowed to participate once). During the HIT, workers 
were presented images one at a time, and, crucially, there was no distinction
or interruption between the initial and test tasks.

A $\chi^2$ test (\#) shows that the frequency of word usage by
workers in the \textit{food} and \textit{objects} treatments differed
significantly during test tasks, even though the test tasks were identical 
for the treatments.  This shows that earlier tasks have a 
significant effect on later ones, and raises concerns about the 
design of microtasks, and possibly the microtask methodology itself.

While this affirms the \textit{existence} of intertask effects, we wish to
understand the nature and severity of the bias induced by prior task 
exposure.  To formalize a notion of the extent of bias, we must first 
acknowledge that workers' responses to microtasks
carry an element of stochasticity: two workers will not, in general, provide
the same response to a given task.  Because worker responses are stochastic,
we take the extent of bias, $\theta$, to be the total shift in probability 
mass, for giving certain responses (i.e. certain words),
caused by a perturbation like intertask effects (\# in other words, the 
bias is the L1-distance between the probability distributions for responses
under one treatment, to that for responses under another.)

It is difficult to measure the change in a probability distribution based on
samples alone (\#), so we reformulate the problem slightly. Consider that,  
no matter the ultimate purpose of a microtask study, a microtask response
is used in \textit{some kind} of subsequent decision or 
further processing.
We denote this dowstream decision by $\mathcal{D}$.  Since $\mathcal{D}$
takes a microtask response as input, biases in the microtask response will 
in general lead to biases in the output of $\mathcal{D}$.  
We can make use of this fact, 
by placing a machine learning algorithm (a classifier), in the role of 
$\mathcal{D}$.  This classifier takes as input a worker's responses to test 
tasks, from which it learns to infer the worker's treatment 
(i.e. the initial tasks she was given). 
Intuitively, the classifier's ability to infer the worker's treatment
depends on how strongly the exposure to prior tasks influences responses to 
subsequent ones.  Formally, the bias, $\theta$, is bounded 
by the accuracy of the classifier, $\eta$ (\#):
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}
Thus, measuring the performance of the classifier allows us to establish a
lower bound on the extent of bias. This approach can be used to bound the 
bias induced by any purturbation, including both intertask effects and 
framing.  Bias ranges from 0 (meaning there are no intertask effects), to 1 
(or 100\%) (meaning that changing initial tasks leads to a completely 
different set of responses to test tasks).

\begin{figure}
	\centering
	\includegraphics[scale=0.6]{figs/images.pdf}
	\caption{
		Examples of images used for various experiments and treatments:
		(A) initial task for \textit{task1:food}, (B) initial task for 
		\textit{task1:obj}, (C) test task for treatments in \textit{exp1}, 
		(D) initial task for \textit{task2:food}, 
		(E) initial task for \textit{task2:cult},  
		(F) test task for treatments in \textit{exp2}.  
		The full set of experimental materials is shown in the 
		supplementary text.
	}

@@

%\setlength{\tabcolsep}{2pt}
%\begin{table}[t]
%\centering
%\begin{tabular}{ c c c c c }
%		\hline \noalign{\smallskip}
%		\multicolumn{3}{c}{\textbf{Treatment name}} & \parbox[c]{1.4cm}{\centering \textbf{Frame}} & \parbox[c]{1.3cm}{\centering \textbf{Initial\\ tasks}}	\\ 
%
%		\noalign{\smallskip} \hline \noalign{\smallskip}
%
%		\multirow{6}{*}{ \parbox[c]{0.8cm}{ \phantom{XX} exp1}} 
%			& \multirow{2}{*}{task1} & food & none & food\\
%			& & obj & none & objects\\
%
%			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
%			& \multirow{2}{*}{frame1} & food & food & none\\
%			& & obj & objects & none\\
%
%			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
%			& \multirow{2}{*}{echo1} & food & food & none\\
%			& & obj & objects & none\\
%
%		\noalign{\smallskip} \hline \noalign{\smallskip}
%
%		\multirow{4}{*}{\parbox[c]{0.8cm}{exp2}} 
%			& \multirow{2}{*}{task2}  &  food & none & food\\
%			& 	&  cult & none & culture\\
%			\noalign{\smallskip} \cline{2-5} \noalign{\smallskip}
%			& \multirow{2}{*}{frame2} & food & food & food\\
%			& 	& cult & culture & food\\
%
%		\noalign{\smallskip} \hline  
%	\end{tabular}
%
%	\caption{ 
%		We use the treatment names shown to refer to specific sections of 
%		the experimental data.
%	}
%	\label{table:treatments}
%\end{table}

Using a na\"ive Bayes classifier (\# in the 
supplementary material we discuss the choice of classifier, the
approach to measuring accuracy, and draw a comparison to the results obtained 
if a support vector machine is used.), we found a bias of 30\% between 
workers from the \textit{food} and \textit{objects} treatments, due to 
intertask effects.
As a point of comparison, we performed a similar experiment in which we 
framed the purpose of the tasks.  Workers were either told that the tasks 
were ``Funded by the laboratory for the visual perception of Food and 
Ingredients'', or ``\ldots of Objects and Tools''.  We estimated the bias 
induced by framing using a na\"ive Bayes classifier, but did not find a 
statistically significant effect (see Fig.~\ref{fig:theta}).  This suggests
that prior task exposure is a stronger priming modality than framing.

We conducted variants of these experiments using different images, to see 
whether this trend was robust.  In the experiment 
\textit{intertask-food-culture},
workers were either assigned to a \textit{food} or \textit{culture} treatment.
As before, the initial tasks for the \textit{food} treatment contained images 
depicting food (\#) (see Fig.~\ref{fig:task}D). 
In the \textit{culture} treatment, 
the initial tasks contained images depicting cultural scenes 
(of dance, sport, or music) (see Fig.~\ref{fig:task}E).  The test tasks, which were identical for both
treatments, included images depicting meals of identifiable cultural 
origin 
(see Fig.~\ref{fig:task}F).  Results for this experiment again showed a 
strong bias (about 50\%; see Fig.~\ref{fig:theta}A) between the labels 
provided by workers from the two 
treatmens during test tasks.  Again, we compared intertask effects to framing,
by performing an experiment (\textit{frame-food-culture}) using 
the same test tasks as in \textit{intertask-food-culture}, in which we exposed
workers to a statement framing the purpose of the tasks as the recognition 
of either food or cultures.  
For this framing experiment, we did not observe a statistically significant 
difference between the labels attributed by the \textit{food} and 
\textit{culture} treatments during test tasks (see Fig.~\ref{fig:theta}).

It was only when we combined the use of framing, with an active 
reiteration step, that bias reached a level comparable to that achieved 
by intertask effects.  In the experiment \textit{echo-food-objects},
after framing the purpose of the task around either the recognition of food
or objects, we required the worker to echo back the purpose of the task
using combo-box input.  In this case, workers exposed to different 
``echoed framing'' showed a bias of about 35\% in the labels provided during 
test tasks
(see Fig.~\ref{fig:theta}A). However, it is possible that by requiring 
workers to reiterate the language used in framing, we provided an implicit 
but nonetheless strong signal that the worker must act on the framing 
statement.  In this sense, \textit{echo-food-culture} may reflect the 
influence of an instruction rather than framing.

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Bias as measured by a na\"ive Bayes classifier, $\theta_\mathrm{NB}$,
		induced in image-labeling tasks, due to framing and 
		intertask effects. 
		(A) Bias in the responses to five test tasks,
		measured between the pairs of treatments indicated on the abscissa.  
		(B) Bias between the treatments of \textit{task1}, 
		measured for individual test tasks, and plotted as a function of 
		task position.  Error bars show the 95\% confidence intervals.
	}
	\label{fig:theta}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/vocab_specificity.pdf}
	\caption{
		Intertask and framing effects on the vocabulary workers use to refer 
		to food(\#), when labeling images, for various treatment pairs.
		(A) Relative change in the number references to food. 
		(B) Relative change in the number of \textit{unique} 
		references to food (lexical richness). 
		(C) Relative specialization of food references, based on the 
		hypernym-hyponym relations (this calculation is shown in the 
		supplementary text).  For all
		plots, a positive bar height indicates that the \textit{*:food}
		treatment has more of the quantity than its conjugate treatment.
	}
	\label{fig:specificity}
\end{figure}

To better understand the nature of the 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. A natural expectation is that, within
a given experiment, the food-primed workers would use more food-related words 
than their non-food-primed counterparts.  However, this was not generally 
the case. In the \textit{intertask-food-culture} experiment, food-primed
workers actually used significantly \textit{fewer} food-related words 
during test tasks
(\# food related words were identified using the wordnet corpus, augmented
with additional words obtained by crawling a recipe website (see supplementary
information for details)) (see Fig.~\ref{fig:specificity}).  This finding
rules out a simplistic conception of how intertask effects guide the worker's
attention.  It seems that when workers encounter content in 
prior tasks, the probability that they will refer to similar content in 
test tasks changes significantly, although it does not necessarily increase.

To clarify this observation, we investigated workers' lexical richness in 
reference to food, that is, the number of \textit{unique} food-related words
used.
Both \textit{intertask} experiments showed that food-exposed workers had 
greater lexical richness, in reference to food, than their non-food-primed 
counterparts (as much as 20\% more; see Fig.~\ref{fig:specificity}).  This
held in the case of \textit{intertask-food-culture}, even though, in that 
experiment, food-primed workers referred to food less often overall.  
This enrichment of the food-related lexicon 
was also observed in the \textit{priming-food-objects}, and 
\textit{echo-food-objects} experiments, but to a lesser extent.

This suggests that priming, by prior task exposure as well as framing,
might have led workers to use more refined or specialized words to refer to
the concept present in the prime.
To test this directly, we used the wordnet corpus to operationalize the 
notion of word specialization.
Wordnet provides a set of hypernym-hyponym relations between \# English 
nouns.  
Hypernyms are generalizations (for example, ``bread'' is a hypernym for 
``pumpernickel''), while hyponyms are specializations.
To assess the relative specialization of words from two different treatments,
we compared all pairs of words from one treatment to the other. 
We tallied the cases where a word from one treatment was more specialized
than one from the other, as well as the cases with roles reversed.  
The difference of these tallies, as a percentage of their total, gave the
relative specialization of one treatments vocabulary over the other
(\#) (see Fig.~\ref{fig:specificity}).  
Using this approach, we compared the specialization of food-related words 
for the treatments of all experiments.

In all experiments except \textit{frame-food-culture}, food-primed workers 
used significantly more specialized words than their non-food-primed 
counterparts (about 15\% more; see Fig.~\ref{fig:specificity}C).
It is interesting that such substantial increases in both the lexical 
richness and specialization of food-related words among food-primed workers
held for \textit{intertask-food-culture}, where, again, we observed fewer 
references to food by food-primed workers overall. These observations point 
to countervailing factors: one tending to activate the more specialized and 
less common food-related words (yielding greater lexical richness and 
specialization), and the other tending to suppress certain, presumably more 
common or generic words (yielding fewer food-related words in total).

This hypothesis is corroborated when we look at those words whose 
frequencies changed the most from one treatment to another (see Table S1).  
The word ``food'', which is the most generic possible food-related word, was 
always \textit{suppressed} among food-primed workers.  In fact, 
for all experiments, ``food'' was the \textit{most suppressed} word.

Taken together, our results can be 
explained through a combination of positive and negative priming.
Positive priming (usually simply ``priming'') occurs when a prior stimulus 
predisposes a person to give certain responses in an ensuing task, and
is often observed as an increase in the speed or accuracy of a response, or
the ability to recognize features of a briefer, quieter, or noisier stimulus 
(\#).
Negative priming occurs when, after being repeatedly exposed to a stimulus 
considered to be non-salient, a person begins to ignore that 
stimulus (\#).

Workers exposed to images containing food will be (positively) primed, 
activating memories, concepts, and vocabulary related to food.  
However, as the worker labels successive images containing food, the basic 
fact that an image contains food will not appear to be salient, since it 
does not serve to distinguish one image from another.  Thus, the most 
generic references to that fact, such as the label ``food'', 
will be suppressed, while more specialized references will be elicited.  
Meanwhile, the number of references to food overall might increase or 
decrease, depending on the balance of these factors.

More generally, we are suggesting that, even though workers are not 
instructed to compare tasks in any way, prior tasks form a 
context relative to which workers judge salience.  Thus, due to a combination 
of negative and positive priming, a worker's focus 
in repeated tasks tends to be directed away from generic, shared features, 
toward specific and distinguishing ones.

Prior to this study, little thought appears to have been given to the 
ordering and bundling of tasks.  But our findings show that these are 
powerful design factors.  A natural response to our findings 
might be simply to randomize task ordering, a practice that is commonly 
observed.  But the sheer extent of bias that we observed leads us to question 
whether that is the best approach, since this would tend to raise 
the ``noise floor'' of the study.

Furthermore, our findings suggest that, through careful task engineering,
it might be able to achieve much more.  The boost that we observed to the 
specialization and richness of workers' responses motivates further 
investigation to acquire a deeper understanding of intertask effects.  
We hope that future work will yield techniques to leverage intertask effects, 
enabling one to tune the focus, diversity, and specialization of worker 
responses.

A consistent goal in human computation is the achievement of expert-level
judgments from non-expert workers.  This has been achieved in some
applications (\#). The distinction between experts and novices can be 
attributed in part to the knowledge of facts and 
heuristics. But to a large degree, this distinction lies in the ability of 
experts to direct their focus towards salient features, while filtering out 
non-salient ones \cite{kellman2009perceptual}.  Using prior task exposure to 
guide focus and salience attribution might well enable expert-level 
judgment in a wider variety of crowdsourced applications.

Here we have shown that a large bias can be introduced into microtask 
responses, simply from the performance of earlier tasks. 
These effects are much stronger than framing, and are only comparable when
framing is combined with a requirement that the worker actively echo
the framing language.  While our findings raise serious concerns about the 
current state of microtask design, it also reveals opportunities for more 
refined control over worker focus and acuity.  Any application that relies
on the repetitive use of human judgment is likely subject to the phenomena
we described here.  It remains to be determined whether intertask effects 
can be leveraged to increase the performance and reliability of human 
computation.
